---
title: "Assignment - Min Jeong"
---
```{r setup, include=FALSE}
library(tidycensus)
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(readr)

options(tigris_use_cache = TRUE)
```

### 1. Data and Variables
```{r, fig.height= 5, fig.width= 5, fig.align='center'}
vars <- load_variables(2024, "acs5")

vars |>
  dplyr::filter(grepl("^B15003", name)) |>
  dplyr::select(name, label, concept) |>
  print(n = 10)

state_abbr <- "TX"
geo_level <- "county" # options: state, county, tract, block group
my_vars <- c(
            income = "B19013_001",
            bachelor = "B15003_022",
            master = "B15003_023",
            professional = "B15003_024",
            doctorate = "B15003_025",
            eduden = "B15003_001")

year_acs <- 2024
survey <- "acs5"    

acs_wide <- get_acs(
            geography = geo_level,
            variables = my_vars,
            state = state_abbr,
            year = year_acs,
            survey = survey,
            geometry = TRUE,
            output = "wide"
            )

acs_wide <- acs_wide |>
  mutate(
    eduE = bachelorE + masterE + professionalE + doctorateE,
    
    eduM = sqrt(
      bachelorM^2 +
      masterM^2 +
      professionalM^2 +
      doctorateM^2
    )
  )

acs_wide <- acs_wide |>
  mutate(
    edu_rate = eduE / edudenE,
    
    edu_moe = moe_prop(
      num = eduE,
      denom = edudenE,
      moe_num = eduM,
      moe_denom = edudenM
    )
  )
```

### 2. Median Household Income
```{r, fig.height= 5, fig.width= 5, fig.align='center'}
ggplot(acs_wide) +
  geom_sf(aes(fill = incomeE), color = NA) +
  scale_fill_viridis_c(name = "Median HH Income") +
  labs(
    title = paste0("ACS ", year_acs,
                   " 5-year: Median Household Income — ",
                   state_abbr, " (", geo_level, ")"),
    caption = "Source: U.S. Census Bureau via tidycensus"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.caption = element_text(size = 8),
    legend.position = "inside",
    legend.position.inside = c(0.2, 0.15)
  )
```

### 3. Educational Attainment
```{r, fig.height= 5, fig.width= 5, fig.align='center'}
top10 <- acs_wide |>
  st_drop_geometry() |>
  arrange(desc(edu_rate)) |>
  select(NAME, edu_rate, edu_moe) |>
  slice_head(n = 10) |>
  mutate(
    edu_rate = edu_rate * 100,
    edu_moe = edu_moe * 100
  )

bottom10 <- acs_wide |>
  st_drop_geometry() |>
  arrange(edu_rate) |>
  select(NAME, edu_rate, edu_moe) |>
  slice_head(n = 10) |>
  mutate(
    edu_rate = edu_rate * 100,
    edu_moe = edu_moe * 100
  )

knitr::kable(
  top10,
  digits = 2,
  row.names = FALSE,
  col.names = c("County", "Bachelor's Degree or Higher (%)", "MOE"),
  caption = "Top 10 Texas Counties by Bachelor's Degree or Higher"
)

knitr::kable(
  bottom10,
  digits = 2,
  row.names = FALSE,
  col.names = c("County", "Bachelor's Degree or Higher (%)", "MOE"),
  caption = "Bottom 10 Texas Counties by Bachelor's Degree or Higher"
)

write_csv(st_drop_geometry(acs_wide),
paste0("acs_", state_abbr, "_", year_acs, ".csv"))
```

### 4. Uncertainty Check

The confidence intervals for Travis County (55.88%–57.24%) and Collin County (55.47%–56.96%) overlap. Although Travis County has a slightly higher estimated rate of bachelor's degree attainment or higher, the overlapping intervals suggest that the difference between the two counties is uncertain. Therefore, their ranking should not be interpreted as a definitive difference in educational attainment.

### 5. Methods Note
This analysis uses the 2024 ACS 5-year estimates for counties in Texas. Median household income (`B19013_001`) was selected for the choropleth map. Educational attainment was measured as the percentage of the population age 25 and older with a bachelor's degree or higher. This percentage was calculated by combining the ACS estimates for bachelor's, master's, professional, and doctorate degrees (`B15003_022`–`B15003_025`) and dividing the sum by the total population age 25 and older (`B15003_001`). Margins of error were carried through the calculations rather than treating the ACS estimates as exact values. The MOEs show that estimates for some counties, especially counties with smaller populations, are much less precise. Therefore, differences in county rankings should be interpreted cautiously when their confidence intervals overlap.